\chapter*{Acknowledgements}

\addcontentsline{toc}{chapter}{Acknowledgements}
\vspace{-1cm}

I would like to express my sheer gratitude to the people who have not only made the writing of this thesis possible but also turned it into an extremely interesting and pleasant experience.

First of all, I would like to thank my supervisors Fabrice Balli and Maarten Boonekamp for their constant support, friendly attitude, and endless patience. During the three years of the doctorate, Fabrice and Maarten have spent many hours of their priceless time in enlightening discussions with the thesis author. It was a joy and an honour for me to have been supervised by these exceptionally nice and intelligent people.

My deepest thanks to the jury members: president of the jury David Rousseau, jury member Aram Apyan and thesis reviewers Alessandro Vicini and Andrew Pilkington. Your profound questions and keen comments have helped to greatly improve this thesis.

It was great to be a part of the DPhP division of CEA Saclay and ATLAS group in particular. I would like to thank ATLAS group leaders Claude Guyot and Frederic Deliot, DPhP directors Gauthier Hamel de Monchenault and Nathalie Besson, vice director Georges Vasseur and direction assistant Martine Oger.

I was extremely lucky to work with an amazing analysis team: Luis Fayard, Jan Kretschmar, Ludovica Aperio Bella, Hicham Atmani, thank you. Special cheers to my beloved office-mates and friends, Nancy and Artur. Nancy, you are the kindest person I have ever encountered. Artur, I am so thankful for your amazing sense of humour, even when things were hard.

It was an invaluable experience to make friends during the M2 NPAC study and I am very happy to have you in my life. Luis, my best binome, I believe there is no such physics problem that we would not be able to crack by just discussing it. Jad, Konie, Manuel, Tetiana, our ATLAS journey was full of fun - and more is to come. Tony, you have become my friend at the moment when I needed it most and I will never forget that moment.

I would like to thank Volodymyr Skalozub, my first supervisor and extraordinary physicist who has helped me to find my way to physics. Oleksii Gulov, thank you for becoming my friend after being a great teacher and mentor. My fellow physicists, colleagues in translation, and dear friends Andrii and Ivan, you are the best team I have ever had.

Last but definitely not least sincere thanks to my family for all the love and support, which was vital. Dad, you were the first physicist I have met in my life and my first inspiration. Mom, you have shown me that the world is big and it is worth exploring. My dear brother Ilya, you are my best friend since I was born. My darling fiancée Natalie, thank you for your love, kindness, and patience. We have made it together.